---
title: "Course Project 1"
author: "Theres Thomas"
date: "05/04/2020"
output: html_document
---


# This is my R Markdown Assignment
```{r initial setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##   Loading and preprocessing the data
```{r}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", "activity_data.zip")
unzip("activity_data.zip")
act <- read.csv("activity.csv", stringsAsFactors = F)
```

Install ggplot2 library
```{r ggplot2, echo = FALSE}
# install lubridate for date analysis
x = 0
tryCatch( expr = {
        library(ggplot2)
        x = 1
    }
)
if (x == 0) {
    install.packages("ggplot2")
    library(ggplot2)
}
```

Convert the dates to date format
```{r}
act$date <- as.Date(act$date)
```


## What is mean total number of steps taken per day ?

### Plotting the Histogram
```{r}
agg0 <- aggregate(steps ~ date, FUN = sum, data = act)
qplot(agg0$steps)
```

### Finding mean and median

*Mean* of steps taken each day
```{r}
(agg1 <- aggregate(steps ~ date, FUN = mean, data = act))
```

*Median* number of steps taken each day
```{r}
(agg2 <- aggregate(steps ~ date, FUN = median, data = act))
```

<br><br>

## What is the average daily activity pattern ?

### Time series plot
```{r}
qplot(agg1$date, agg1$steps) +
    geom_line(aes(x = agg1$date, y = agg1$steps), colour = "blue")
```

### The 5-minute interval contains the max number of steps
```{r}
loc <- which(act$steps == max(na.omit(act$steps)))
act[loc,]
```



##  Imputing missing values

Here *mean imputation* is used
```{r}
avg <- mean(na.omit(act$steps))
avg <- floor(avg)  # round down
```

### assigning the avg value to all NA in a new dataset:
```{r}
imputed <- act
for (i in 1:length(imputed$steps)) {
    if (is.na(imputed$steps[i])) {
        imputed$steps[i] <- avg
    }
}
```

### Histogram after missing values are imputed:
```{r}
agg3 <- aggregate(steps ~ date, FUN = sum, data = act)
qplot(agg3$steps)
```

<br><br>

## Differences in activity patterns between weekdays and weekends

```{r}
loc <- which(weekdays(imputed$date) %in% c("Saturday", "Sunday"))
weekday <- imputed[-loc,]
weekend <- imputed[loc,]
agg4 <- aggregate(steps ~ interval, data = weekday, FUN = mean)
agg5 <- aggregate(steps ~ interval, data = weekend, FUN = mean)
agg4$w <- rep("weekday", length(agg4[,1]))
agg5$w <- rep("weekend", length(agg5[,1]))
agg4 <- rbind(agg4, agg5)
ggplot(data = agg4, mapping = aes(x = interval, y = steps)) +
    geom_line() + facet_wrap(~w, nrow = 2)
```